---
title: "BeaverNutrientAnalysis"
format: html
editor: visual
---

# Beaver Dam Nutrient Analysis

**Purpose** The purpose of this code is to examine patterns in nutrients (NOx, PO4 and NH4), temperature, and specific conductivity above and below beaver dams in Dry Creek Experimental Watershed, Boise, ID.

**Output** The output of this code is two box and whisker figures. The first groups samples by above and below a beaver dam at two separate sites, a lower and upper dam. One plot analyzes each sample day separately, while the other combines values measured across all days.

**To do** Increase font size

## Load Libraries

```{r}
#| warning: false
library(ggplot2)
library(readxl)
library(dplyr)
library(cowplot)
```

## Bring in data

```{r}
setwd("/Users/hannahrichardson/Documents/BSU/Spring '23/Field Methods/FieldMethodsHydro")
list.files()
df <- read.csv("Quality Beaver Meadows Data Template.csv")
```

## Average data from all weeks

```{r}
#Create above and below pool box and whisker plots for each measured value
NOx <- df %>%
  ggplot(aes(x = Site_Location, y = NOx_mgL, fill= Site_type))+
  geom_boxplot(show.legend = FALSE)+
  xlab(NULL)+
  ylab('NOx (mg/L)')+
  scale_y_continuous(n.breaks=7)+
  theme_bw()+
  theme(text = element_text(size = 14))  

PO4 <- df %>%
  ggplot(aes(x = Site_Location, y = PO4_mgL, fill= Site_type))+
  geom_boxplot(show.legend = FALSE)+
  xlab(NULL)+
  ylab('PO4 (mg/L)')+
  scale_y_continuous(n.breaks=7)+
  theme_bw()+
  theme(text = element_text(size = 14))  

NH3 <- df %>%
  ggplot(aes(x = Site_Location, y = NH4_mgL, fill= Site_type))+
  geom_boxplot(show.legend = FALSE)+
  xlab(NULL)+
  ylab('NH3 (mg/L)')+
  scale_y_continuous(n.breaks=7)+
  theme_bw()+
  theme(text = element_text(size = 14))  

temp <- df %>%
  ggplot(aes(x = Site_Location, y = Temp._C, fill= Site_type))+
  geom_boxplot()+
  xlab(NULL)+
  ylab('Temp C')+
  theme_bw()+
      theme(legend.position = c(0.25, 0.73),
      legend.background = element_rect( fill = alpha("white", 0)))+
    labs(fill = NULL)+
    scale_y_continuous(n.breaks=7)+
  theme(text = element_text(size = 14))  

EC <- df %>%
  ggplot(aes(x = Site_Location, y = EC, fill= Site_type))+
  geom_boxplot(show.legend = FALSE)+
  xlab('Site Type')+
  ylab('EC (uS/cm)')+
  scale_y_continuous(n.breaks=7)+
  theme_bw()+
  theme(text = element_text(size = 14))  

#combine each individual plot into a column 
Nutrient_plots<-plot_grid(NOx, PO4, NH3, temp, EC, 
                nrow =5)
```

## Save pdfs

```{r}
pdf(file = "Weekscombined.pdf",
    width = 4,
    height = 10)

Nutrient_plots

dev.off()
```
