---
title: "BeaverNutrientAnalysis"
format: html
editor: visual
---

# Beaver Dam Nutrient Analysis

**Purpose** The purpose of this code is to examine patterns in nutrients (NOx, PO4 and NH3), temperature, and specific conductivity above and below beaver dams in Dry Creek Experimental Watershed, Boise, ID.

**Output** The output of this code is two box and whisker figures. The first groups samples by above and below a beaver dam at two separate sites, a lower and upper dam. One plot analyzes each sample day separately, while the other combines values measured across all days.

**To do** standardize y axis range

## Load Libraries

```{r}
#| warning: false
library(ggplot2)
library(readxl)
library(dplyr)
library(cowplot)
```

## Bring in data

```{r}
setwd("/Users/hannahrichardson/Documents/BSU/Spring '23/Field Methods/FieldMethodsHydro")
list.files()
df <- read.csv("Quality Beaver Meadows Data Template.csv")
```

```{r}
#set consistent y limits for each plot

noxlim<-c(0.032,0.044)
po4lim<-c(0.016,0.024)
nh4lim<-c(0.0150,0.0225)
templim<-c(0,8)
eclim<-c(130,190)
```

## Week 1

Pulls out data from Week 1 of analysis and builds box and whisker plots

```{r}
#Set date
date<- "2/24/23"

#Create above and below pool box and whisker plots for each measured value
NOx <- df %>%
  filter(Date == date)%>%
  ggplot(aes(x = Site_type, y = NOx_mgL, fill= Site_Location))+
  geom_boxplot(show.legend = FALSE)+
  xlab(NULL)+
  ylab('NOx (mg/L)')+
  ggtitle('Week 1')+
  theme(plot.title = element_text(hjust = 0.5))+
  ylim(noxlim)+
  theme_bw()

PO4 <- df %>%
  filter(Date == date)%>%
  ggplot(aes(x = Site_type, y = PO4_mgL, fill= Site_Location))+
  geom_boxplot(show.legend = FALSE)+
  xlab(NULL)+
  ylab('PO4 (mg/L)')+
  ylim(po4lim)+
  theme_bw()

NH4 <- df %>%
  filter(Date == date)%>%
  ggplot(aes(x = Site_type, y = NH4_mgL, fill= Site_Location))+
  geom_boxplot(show.legend = FALSE)+
  xlab(NULL)+
  ylab('NH4 (mg/L)')+
  ylim(nh4lim)+
  theme_bw()

temp <- df %>%
  filter(Date == date)%>%
  ggplot(aes(x = Site_type, y = Temp._C, fill= Site_Location))+
  geom_boxplot(show.legend = FALSE)+
  xlab(NULL)+
  xlab('Site Type')+
  ylab('Temp C')+
  ylim(templim)+
  theme_bw()

EC <- df %>%
  filter(Date == date)%>%
  ggplot(aes(x = Site_type, y = EC, fill= Site_Location))+
  geom_boxplot(show.legend = FALSE)+
  xlab('Site Type')+
  ylab('EC (uS/cm)')+
  ylim(eclim)+
  theme_bw()

#combine each individual plot into a column 
Wk1<-plot_grid(NOx, PO4, NH4, temp, EC, 
                nrow =5)

```

## Week 2

Pulls out data from Week 2 of analysis and builds box and whisker plots

```{r}
#set date
date<- "3/3/23"

#Create above and below pool box and whisker plots for each measured value
NOx <- df %>%
  filter(Date == date)%>%
  ggplot(aes(x = Site_type, y = NOx_mgL, fill= Site_Location))+
  geom_boxplot(show.legend = FALSE)+
  xlab(NULL)+
  ylab(NULL)+
  ggtitle('Week 2')+
  theme(plot.title = element_text(hjust = 0.5))+
  ylim(noxlim)+
  theme_bw()

PO4 <- df %>%
  filter(Date == date)%>%
  ggplot(aes(x = Site_type, y = PO4_mgL, fill= Site_Location))+
  geom_boxplot(show.legend = FALSE)+
  xlab(NULL)+
  ylab(NULL)+
  ylim(po4lim)+
  theme_bw()

NH4 <- df %>%
  filter(Date == date)%>%
  ggplot(aes(x = Site_type, y = NH4_mgL, fill= Site_Location))+
  geom_boxplot(show.legend = FALSE)+
  xlab(NULL)+
  ylab(NULL)+
  ylim(nh4lim)+
  theme_bw()

temp <- df %>%
  filter(Date == date)%>%
  ggplot(aes(x = Site_type, y = Temp._C, fill= Site_Location))+
  geom_boxplot(show.legend = FALSE)+
  xlab(NULL)+
  ylab(NULL)+
  ylim(templim)+
  theme_bw()

EC <- df %>%
  filter(Date == date)%>%
  ggplot(aes(x = Site_type, y = EC, fill= Site_Location))+
  geom_boxplot(show.legend = FALSE)+
  xlab('Site Type')+
  ylab(NULL)+
  ylim(eclim)+
  theme_bw()

#combine each individual plot into a column 
Wk2<-plot_grid(NOx, PO4, NH4, temp, EC, 
                nrow =5)

```

## Week 3

Pulls out data from Week 3 of analysis and builds box and whisker plots

```{r}
#set date
date<- "3/10/23"

#Create above and below pool box and whisker plots for each measured value
NOx <- df %>%
  filter(Date == date)%>%
  ggplot(aes(x = Site_type, y = NOx_mgL, fill= Site_Location))+
  geom_boxplot(show.legend = FALSE)+
  xlab(NULL)+
  ylab(NULL)+
  ggtitle('Week 3')+
  theme(plot.title = element_text(hjust = 0.5))+
  ylim(noxlim)+
  theme_bw()

PO4 <- df %>%
  filter(Date == date)%>%
  ggplot(aes(x = Site_type, y = PO4_mgL, fill= Site_Location))+
  geom_boxplot(show.legend = FALSE)+
  xlab(NULL)+
  ylab(NULL)+
  ylim(po4lim)+
  theme_bw()

NH4 <- df %>%
  filter(Date == date)%>%
  ggplot(aes(x = Site_type, y = NH4_mgL, fill= Site_Location))+
  geom_boxplot(show.legend = FALSE)+
  xlab(NULL)+
  ylab(NULL)+
  ylim(nh4lim)+
  theme_bw()

temp <- df %>%
  filter(Date == date)%>%
  ggplot(aes(x = Site_type, y = Temp._C, fill= Site_Location))+
  geom_boxplot()+
  xlab(NULL)+
  ylab(NULL)+
  theme_bw()+
  ylim(templim)+
  theme(legend.position = c(0.35, 0.7),
        legend.background = element_rect( fill = alpha("white", 0)))+
        labs(fill = "Site Location")

EC <- df %>%
  filter(Date == date)%>%
  ggplot(aes(x = Site_type, y = EC, fill= Site_Location))+
  geom_boxplot(show.legend=FALSE)+
  xlab('Site Type')+
  ylab(NULL)+
  ylim(eclim)+
  theme_bw()

#combine each individual plot into a column 
Wk3<-plot_grid(NOx, PO4, NH4, temp, EC, 
                nrow =5)
```

### Combine weekly plots

Combine box and whisker plots into one figure

```{r}
allplot<-plot_grid(Wk1,Wk2,Wk3,
                   ncol= 3)
```

## Average weeks

```{r}
#Create above and below pool box and whisker plots for each measured value
NOx <- df %>%
  ggplot(aes(x = Site_type, y = NOx_mgL, fill= Site_Location))+
  geom_boxplot(show.legend = FALSE)+
  xlab(NULL)+
  ylab('NOx (mg/L)')+
  theme_bw()

PO4 <- df %>%
  ggplot(aes(x = Site_type, y = PO4_mgL, fill= Site_Location))+
  geom_boxplot(show.legend = FALSE)+
  xlab(NULL)+
  ylab('PO4 (mg/L)')+
  theme_bw()

NH4 <- df %>%
  ggplot(aes(x = Site_type, y = NH4_mgL, fill= Site_Location))+
  geom_boxplot(show.legend = FALSE)+
  xlab(NULL)+
  ylab('NH4 (mg/L)')+
  theme_bw()

temp <- df %>%
  ggplot(aes(x = Site_type, y = Temp._C, fill= Site_Location))+
  geom_boxplot()+
  xlab(NULL)+
  ylab('Temp C')+
  theme_bw()+
      theme(legend.position = c(0.15, 0.7),
      legend.background = element_rect( fill = alpha("white", 0)))+
  labs(fill = "Site Location")

EC <- df %>%
  ggplot(aes(x = Site_type, y = EC, fill= Site_Location))+
  geom_boxplot(show.legend = FALSE)+
  xlab('Site Type')+
  ylab('EC (uS/cm)')+
  theme_bw()

#combine each individual plot into a column 
Combo<-plot_grid(NOx, PO4, NH4, temp, EC, 
                nrow =5)
```

## Save pdfs

```{r}
pdf(file = "Allplots.pdf",
    width = 6,
    height = 10)

allplot
dev.off()

pdf(file = "Weekscombined.pdf",
    width = 4,
    height = 10)

Combo
dev.off()
```
